---
title: Countour plots for each tissue _ normalising data to 2500 cells per tissue
  / Donor
author: "Nicole Mende"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## input: Table with cell scores/pseudotime/UMAP coordinates (columns) for each individual cell (rows) = “Table_Dpt_Cell_scores.csv”

## Load libraries
```{r}
library(ggplot2)
library(grid)
```

## Preparation of data
```{r}
## load data file
data <- read.csv("Table_Dpt_Cell_scores.csv", header = TRUE)

## subset by Donor and tissue
DOD_BM <- subset(data, Tissue=="BM")
DOD_SPL <- subset(data, Tissue=="SPL")
PB <- subset(data, Tissue=="PB")

DOD_BM_d1 <- DOD_BM[(grep("DOD1", DOD_BM$Donor)),]
DOD_BM_d2 <- DOD_BM[(grep("DOD2", DOD_BM$Donor)),]
DOD_BM_d3 <- DOD_BM[(grep("DOD3", DOD_BM$Donor)),]
DOD_BM_d4 <- DOD_BM[(grep("DOD4", DOD_BM$Donor)),]

DOD_SPL_d1 <- DOD_SPL[(grep("DOD1", DOD_SPL$Donor)),]
DOD_SPL_d2 <- DOD_SPL[(grep("DOD2", DOD_SPL$Donor)),]
DOD_SPL_d4 <- DOD_SPL[(grep("DOD4", DOD_SPL$Donor)),]

DOD_PB_d1 <- PB[(grep("DOD1", PB$Donor)),]
DOD_PB_d2 <- PB[(grep("DOD2", PB$Donor)),]

LD_PB_BP1c <- PB[(grep("BP1c", PB$Donor)),]
LD_PB_BP37d <- PB[(grep("BP37d", PB$Donor)),]
LD_PB_BP59h <- PB[(grep("BP59h", PB$Donor)),]
LD_PB_BP62j <- PB[(grep("BP62j", PB$Donor)),]
LD_PB_BP74 <- PB[(grep("BP74", PB$Donor)),]
LD_PB_TQ198 <- PB[(grep("TQ198", PB$Donor)),]
```



## Randomly select 2500 cells from each Donor / Tissue before combining all samples of the same tissues (DODBM, DODPSL, DODPB, LDPB)
```{r}
## BM
R1 <- sample(row.names(DOD_BM_d1),size=2500)
DOD_BM_d1_R <- DOD_BM_d1[c(R1),]

R1 <- sample(row.names(DOD_BM_d2),size=2500)
DOD_BM_d2_R <- DOD_BM_d2[c(R1),]

R1 <- sample(row.names(DOD_BM_d3),size=2500)
DOD_BM_d3_R <- DOD_BM_d3[c(R1),]

R1 <- sample(row.names(DOD_BM_d4),size=2500)
DOD_BM_d4_R <- DOD_BM_d4[c(R1),]

DOD_BM_R <- rbind(DOD_BM_d1_R, DOD_BM_d2_R,DOD_BM_d3_R, DOD_BM_d4_R)


## SPL
R1 <- sample(row.names(DOD_SPL_d1),size=2500)
DOD_SPL_d1_R <- DOD_SPL_d1[c(R1),]

R1 <- sample(row.names(DOD_SPL_d2),size=2500)
DOD_SPL_d2_R <- DOD_SPL_d2[c(R1),]

R1 <- sample(row.names(DOD_SPL_d4),size=2500)
DOD_SPL_d4_R <- DOD_SPL_d4[c(R1),]

DOD_SPL_R <- rbind(DOD_SPL_d1_R, DOD_SPL_d2_R,DOD_SPL_d4_R)


## LD-PB
R1 <- sample(row.names(DOD_PB_d1),size=2500)
DOD_PB_d1_R <- DOD_PB_d1[c(R1),]

R1 <- sample(row.names(DOD_PB_d2),size=2500)
DOD_PB_d2_R <- DOD_PB_d2[c(R1),]

DOD_PB_R <- rbind(DOD_PB_d1_R, DOD_PB_d2_R)


## LD-PB
R1 <- sample(row.names(LD_PB_BP1c),size=2500)
LD_PB_d1_R <- LD_PB_BP1c[c(R1),]

R1 <- sample(row.names(LD_PB_BP37d),size=2500)
LD_PB_d2_R <- LD_PB_BP37d[c(R1),]

R1 <- sample(row.names(LD_PB_BP62j),size=2500)
LD_PB_d3_R <- LD_PB_BP62j[c(R1),]

R1 <- sample(row.names(LD_PB_BP74),size=2500)
LD_PB_d4_R <- LD_PB_BP74[c(R1),]

R1 <- sample(row.names(LD_PB_TQ198),size=2500)
LD_PB_d5_R <- LD_PB_TQ198[c(R1),]

R1 <- sample(row.names(LD_PB_BP59h),size=2500)
LD_PB_d6_R <- LD_PB_BP59h[c(R1),]

LD_PB_R <- rbind(LD_PB_d1_R, LD_PB_d2_R,LD_PB_d3_R, LD_PB_d4_R,LD_PB_d5_R, LD_PB_d6_R)
```



## Plot contour for each tissue with limit max=0.04, bins = 10 (Figure 2a - code will generate slightly different contour plots each time as it uses ransomly selected cells)
```{r}
commonTheme = list(labs(color="Density",fill="Density",
                        x="-UMAP2",
                        y="UMAP1"),
                   theme_bw(),
                  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
                  )

#BM
BM_plot <- ggplot(data=DOD_BM_R,aes(-UMAP2,UMAP1)) + geom_point(cex=0.2, colour = "#e3e3e3") + xlim(-15,15) + ylim (-10,10) + stat_density2d(aes(fill=..level..,alpha=..level..), geom='polygon',colour='black', cex=0.2, bins=10) + 
  scale_fill_continuous(low="#b8b1cc",high="#2a1142", limits = c(0, 0.04), breaks = c(0, 0.015, 0.025, 0.035)) +
  guides(alpha="none") + commonTheme +   theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + theme(legend.title=element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA), axis.title = element_blank())

BM_plot + theme(aspect.ratio=1) 
#ggsave("Contour_BM.png", bg="transparent", dpi = 200)


#SPL
SPL_plot <- ggplot(data=DOD_SPL_R,aes(-UMAP2,UMAP1)) + geom_point(cex=0.2, colour = "#e3e3e3") + xlim(-15,15) + ylim (-10,10) + stat_density2d(aes(fill=..level..,alpha=..level..), geom='polygon',colour='black', cex=0.2, bins=10) + 
  scale_fill_continuous(low="#b8b1cc",high="#2a1142", limits = c(0, 0.04), breaks = c(0, 0.015, 0.025, 0.035)) +
  guides(alpha="none") + commonTheme +   theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + theme(legend.title=element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA), axis.title = element_blank())

SPL_plot + theme(aspect.ratio=1) 
#ggsave("Contour_SPL.png", bg="transparent", dpi = 200)


#DODPB
DODPB_plot <- ggplot(data=DOD_PB_R,aes(-UMAP2,UMAP1)) + geom_point(cex=0.2, colour = "#e3e3e3") + xlim(-15,15) + ylim (-10,10) + stat_density2d(aes(fill=..level..,alpha=..level..), geom='polygon',colour='black', cex=0.2, bins=10) + 
  scale_fill_continuous(low="#b8b1cc",high="#2a1142", limits = c(0, 0.04), breaks = c(0, 0.015, 0.025, 0.035)) +
  guides(alpha="none") + commonTheme +   theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + theme(legend.title=element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA), axis.title = element_blank())

DODPB_plot + theme(aspect.ratio=1) 
#ggsave("Contour_DODPB.png", bg="transparent", dpi = 200)


#LDPB
LDPB_plot <- ggplot(data=LD_PB_R,aes(-UMAP2,UMAP1)) + geom_point(cex=0.2, colour = "#e3e3e3") + xlim(-15,15) + ylim (-10,10) + stat_density2d(aes(fill=..level..,alpha=..level..), geom='polygon',colour='black', cex=0.2, bins=10) + 
  scale_fill_continuous(low="#b8b1cc",high="#2a1142", limits = c(0, 0.04), breaks = c(0, 0.015, 0.025, 0.035)) +
  guides(alpha="none") + commonTheme +  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + theme(legend.title=element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA), axis.title = element_blank())

LDPB_plot + theme(aspect.ratio=1) 
#ggsave("Contour_LDPB.png", bg="transparent", dpi = 200)
```
